---
title: "Initial Epi Data Quality Review"
author: "bcapture"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    self-contained: true
    code-fold: false
css: assets/epi-report.css
execute:
  echo: false
  warning: false
  message: false
params:
  data_file: null
  report_title: "Initial Epi Data Quality Review"
  generated_at: null
---

```{r}
data <- readRDS(params$data_file)
p <- data
manifest <- p$summary_manifest
overview <- p$dataset_overview
status <- as.character(manifest$status[[1L]])
privacy <- data$privacy

fmt_integer <- function(x) {
  ifelse(is.na(x), "-", format(as.integer(round(as.numeric(x))), big.mark = ",", trim = TRUE))
}
fmt_decimal <- function(x, digits = 1L) {
  ifelse(is.na(x), "-", format(round(as.numeric(x), digits), nsmall = 0L,
    trim = TRUE, scientific = FALSE))
}
fmt_percent <- function(x) fmt_decimal(x, 1L)
humanize_name <- function(x) {
  x <- gsub("_+", " ", as.character(x))
  tools::toTitleCase(x)
}
question_label <- function(x) {
  text <- if ("question_text" %in% names(x)) trimws(as.character(x$question_text)) else rep("", nrow(x))
  fallback <- if ("canonical_name" %in% names(x)) humanize_name(x$canonical_name) else rep("", nrow(x))
  use_text <- !is.na(text) & nzchar(text) & nchar(text) <= 80L
  ifelse(use_text, text, fallback)
}
section_label <- function(x) {
  text <- if ("section_name" %in% names(x)) trimws(as.character(x$section_name)) else rep("", nrow(x))
  fallback <- if ("section_id" %in% names(x)) as.character(x$section_id) else rep("", nrow(x))
  ifelse(!is.na(text) & nzchar(text), text, fallback)
}
table_html <- function(x, class = "epi-table") {
  if (nrow(x) == 0L) return("No rows to display.")
  classes <- unique(c("epi-table", strsplit(class, " +")[[1L]]))
  html <- knitr::kable(x, format = "html", escape = TRUE,
    table.attr = paste0('class="', paste(classes, collapse = " "), '"'))
  paste0('<div class="epi-table-scroll">', html, "</div>")
}
fmt <- function(x) ifelse(is.na(x) | !nzchar(as.character(x)), "-", as.character(x))
```

::: {.callout-note}
**`r params$report_title`**

This report describes validation, completeness, parsing, and analytical
readiness for de-identified Initial Epidemiological Interview summary products
for controlled analytical use.
It does not contain the private re-identification crosswalk and is not
automatically suitable for unrestricted public release.
:::

```{r, results='asis'}
if (identical(status, "review")) {
  cat("::: {.callout-warning}\n**Analytical status: REVIEW**\n\nReview findings remain and should be resolved or accepted before downstream use.\n:::\n")
} else {
  cat("::: {.callout-note}\n**Analytical status: PASSED**\n\nThe summary products passed the configured analytical gates.\n:::\n")
}
```

## Analytical readiness

The readiness values below are supplied by `summary_manifest.csv` and
`dataset_overview.csv`; this report does not create a composite quality score.

```{r, results='asis'}
readiness <- data.frame(
  Measure = c("Summary status", "Validation available", "Valid cases", "Review cases",
    "Error cases", "Privacy-excluded fields", "Unsupported retained fields", "Cases"),
  Value = c(status, fmt(manifest$validation_available[[1L]]),
    overview$n_validation_valid[[1L]], overview$n_validation_review[[1L]],
    overview$n_validation_error[[1L]], manifest$n_privacy_excluded_fields[[1L]],
    manifest$n_unsupported_retained_fields[[1L]], manifest$n_cases[[1L]])
)
cat(table_html(readiness))
```

## Validation status

```{r, results='asis'}
validation <- p$validation_status_summary
validation$validation_status <- factor(validation$validation_status,
  levels = c("valid", "review", "error"))
print(ggplot2::ggplot(validation, ggplot2::aes_string(
  x = "validation_status", y = "n_cases", fill = "validation_status"
)) + ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::scale_fill_manual(values = c(valid = "#3b7d5a", review = "#9a6b20", error = "#a94a4a"),
    drop = FALSE) + ggplot2::labs(x = "Validation status", y = "Cases") +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank()))
validation_display <- data.frame(
  Status = as.character(validation$validation_status),
  Cases = fmt_integer(validation$n_cases),
  `Percent of cases` = fmt_percent(validation$percent_of_cases),
  check.names = FALSE
)
cat(table_html(validation_display, "epi-table-compact"))
```

## Validation findings by rule

```{r, results='asis'}
rules <- p$validation_rule_summary
severity_order <- c("ERROR", "WARNING", "INFO")
rules$.severity_order <- match(toupper(as.character(rules$severity)), severity_order)
rules <- rules[order(rules$.severity_order, -as.numeric(rules$n_findings), rules$rule_id), , drop = FALSE]
rules$.severity_order <- NULL
rules_display <- data.frame(
  Severity = as.character(rules$severity),
  Type = humanize_name(rules$validation_type),
  Rule = as.character(rules$rule_id),
  Findings = fmt_integer(rules$n_findings),
  Cases = fmt_integer(rules$n_cases),
  check.names = FALSE
)
cat(table_html(rules_display, "epi-table-compact"))
```

## Numeric parsing quality

Parse-success percentages below are presentation-only ratios of the supplied
`n_parsed` and `n_populated` counts; underlying counts and statistics are not
recomputed.

```{r, results='asis'}
numeric <- p$numeric_summaries
numeric_display <- data.frame(
  Section = section_label(numeric),
  `Question / Variable` = question_label(numeric),
  Units = fmt(numeric$units),
  Entered = fmt_integer(numeric$n_populated),
  Parsed = fmt_integer(numeric$n_parsed),
  Unparsed = fmt_integer(numeric$n_unparsed),
  `Parse success (%)` = fmt_percent(ifelse(numeric$n_populated > 0,
    100 * numeric$n_parsed / numeric$n_populated, NA_real_)),
  check.names = FALSE
)
cat(table_html(numeric_display, "epi-table-wide"))
```

## Date parsing quality

```{r, results='asis'}
dates <- p$date_summaries
date_display <- data.frame(
  Section = section_label(dates),
  `Question / Variable` = question_label(dates),
  Entered = fmt_integer(dates$n_populated),
  Parsed = fmt_integer(dates$n_parsed),
  Unparsed = fmt_integer(dates$n_unparsed),
  `Parse success (%)` = fmt_percent(ifelse(dates$n_populated > 0,
    100 * dates$n_parsed / dates$n_populated, NA_real_)),
  check.names = FALSE
)
cat(table_html(date_display, "epi-table-wide"))
```

## Missingness overview

Missingness percentages below are presentation-only ratios of the supplied
missing and total counts. They are descriptive displays, not quality thresholds.

```{r, results='asis'}
scalar <- p$scalar_frequencies
scalar_missing <- scalar[!duplicated(scalar$canonical_name), , drop = FALSE]
missingness <- dplyr::bind_rows(
  data.frame(Section = section_label(scalar_missing),
    `Question / Variable` = question_label(scalar_missing),
    Total = scalar_missing$n_cases_total, Missing = scalar_missing$n_missing,
    check.names = FALSE),
  data.frame(Section = section_label(numeric),
    `Question / Variable` = question_label(numeric),
    Total = numeric$n_total, Missing = numeric$n_missing,
    check.names = FALSE),
  data.frame(Section = section_label(dates),
    `Question / Variable` = question_label(dates),
    Total = dates$n_total, Missing = dates$n_missing,
    check.names = FALSE)
)
missingness$`Missing (%)` <- ifelse(missingness$Total > 0,
  100 * missingness$Missing / missingness$Total, NA_real_)
missingness <- missingness[order(-missingness$`Missing (%)`, missingness$`Question / Variable`), , drop = FALSE]
missing_display <- missingness
missing_display$Total <- fmt_integer(missing_display$Total)
missing_display$Missing <- fmt_integer(missing_display$Missing)
missing_display$`Missing (%)` <- fmt_percent(missing_display$`Missing (%)`)
shown <- seq_len(min(15L, nrow(missing_display)))
cat(table_html(missing_display[shown, , drop = FALSE], "epi-table-compact"))
if (nrow(missing_display) > length(shown)) {
  cat("<p class='epi-caption'>Showing the 15 variables with greatest missingness; expand for the complete table.</p>")
  cat("<details><summary>Details: complete missingness table</summary>\n")
  cat(table_html(missing_display, "epi-table-wide"))
  cat("</details>\n")
}
```

## Privacy review summary

Only aggregate privacy metadata is included below. Review rows and tested
strings are never staged for the report.

```{r, results='asis'}
if (nrow(privacy$manifest) > 0L) {
  privacy_manifest <- data.frame(
    Status = as.character(privacy$manifest$status),
    Warnings = fmt_integer(privacy$manifest$privacy_warnings),
    Errors = fmt_integer(privacy$manifest$privacy_errors),
    `Free-text fields withheld` = fmt_integer(privacy$manifest$free_text_fields_withheld),
    check.names = FALSE
  )
  cat(table_html(privacy_manifest, "epi-table-compact"), "\n\n", sep = "")
} else
  cat("No de-identification manifest was available.\n\n")
privacy_review <- data.frame(
  `Review records` = fmt_integer(privacy$review$n_review_rows),
  `Cases affected` = fmt_integer(privacy$review$n_cases_affected),
  `Fields affected` = fmt_integer(privacy$review$n_fields_affected),
  check.names = FALSE
)
privacy_audit <- data.frame(
  `Audit findings` = fmt_integer(privacy$leak_audit$n_audit_rows),
  Errors = fmt_integer(privacy$leak_audit$n_errors),
  Warnings = fmt_integer(privacy$leak_audit$n_warnings),
  Status = as.character(privacy$leak_audit$status),
  check.names = FALSE
)
cat(table_html(privacy_review, "epi-table-compact"), "\n\n", sep = "")
cat(table_html(privacy_audit, "epi-table-compact"))
```

## Repeated-table completeness

```{r, results='asis'}
tables <- p$repeated_table_summaries[c("table_name", "n_cases_total",
  "n_cases_with_records", "n_records_total", "median_records_per_case",
  "max_records_per_case")]
names(tables) <- c("Table", "Cases", "Cases with records", "Records",
  "Median / case", "Max / case")
tables$Cases <- fmt_integer(tables$Cases)
tables$`Cases with records` <- fmt_integer(tables$`Cases with records`)
tables$Records <- fmt_integer(tables$Records)
tables$`Median / case` <- fmt_decimal(tables$`Median / case`)
tables$`Max / case` <- fmt_integer(tables$`Max / case`)
cat(table_html(tables, "epi-table-compact"))
```

## Provenance

```{r, results='asis'}
provenance <- data.frame(
  Measure = c("Form version", "Privacy profile", "Summary schema version",
    "Summary status", "Cases", "bcapture package version", "Report generated"),
  Value = c(as.character(manifest$form_version[[1L]]), as.character(manifest$profile[[1L]]),
    as.character(manifest$summary_schema_version[[1L]]), as.character(status),
    as.character(manifest$n_cases[[1L]]), as.character(manifest$package_version[[1L]]),
    as.character(params$generated_at))
)
cat(table_html(provenance))
cat("\n\n<p class='epi-caption'>Generated by bcapture from de-identified Initial Epi summary products.</p>\n")
```
