---
title: "Initial Epi Data Quality Review"
author: "bcapture"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    self-contained: true
    code-fold: false
css: assets/epi-report.css
execute:
  echo: false
  warning: false
  message: false
params:
  data_file: null
  report_title: "Initial Epi Data Quality Review"
  generated_at: null
---

```{r}
data <- readRDS(params$data_file)
p <- data
manifest <- p$summary_manifest
overview <- p$dataset_overview
status <- as.character(manifest$status[[1L]])
privacy <- data$privacy

table_html <- function(x) {
  if (nrow(x) == 0L) return("No rows to display.")
  knitr::kable(x, format = "html", escape = TRUE)
}
fmt <- function(x) ifelse(is.na(x) | !nzchar(as.character(x)), "-", as.character(x))
```

::: {.callout-note}
**`r params$report_title`**

This report describes validation, completeness, parsing, and analytical
readiness for de-identified Initial Epidemiological Interview summary products
for controlled analytical use.
It does not contain the private re-identification crosswalk and is not
automatically suitable for unrestricted public release.
:::

```{r, results='asis'}
if (identical(status, "review")) {
  cat("::: {.callout-warning}\n**Analytical status: REVIEW**\n\nReview findings remain and should be resolved or accepted before downstream use.\n:::\n")
} else {
  cat("::: {.callout-note}\n**Analytical status: PASSED**\n\nThe summary products passed the configured analytical gates.\n:::\n")
}
```

## Analytical readiness

The readiness values below are supplied by `summary_manifest.csv` and
`dataset_overview.csv`; this report does not create a composite quality score.

```{r, results='asis'}
readiness <- data.frame(
  Measure = c("Summary status", "Validation available", "Valid cases", "Review cases",
    "Error cases", "Privacy-excluded fields", "Unsupported retained fields", "Cases"),
  Value = c(status, fmt(manifest$validation_available[[1L]]),
    overview$n_validation_valid[[1L]], overview$n_validation_review[[1L]],
    overview$n_validation_error[[1L]], manifest$n_privacy_excluded_fields[[1L]],
    manifest$n_unsupported_retained_fields[[1L]], manifest$n_cases[[1L]])
)
cat(table_html(readiness))
```

## Validation status

```{r, results='asis'}
validation <- p$validation_status_summary
validation$validation_status <- factor(validation$validation_status,
  levels = c("valid", "review", "error"))
print(ggplot2::ggplot(validation, ggplot2::aes_string(
  x = "validation_status", y = "n_cases", fill = "validation_status"
)) + ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::scale_fill_manual(values = c(valid = "#3b7d5a", review = "#9a6b20", error = "#a94a4a"),
    drop = FALSE) + ggplot2::labs(x = "Validation status", y = "Cases") +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank()))
cat(table_html(validation))
```

## Validation findings by rule

```{r, results='asis'}
rules <- p$validation_rule_summary
severity_order <- c("ERROR", "WARNING", "INFO")
rules$.severity_order <- match(toupper(as.character(rules$severity)), severity_order)
rules <- rules[order(rules$.severity_order, -as.numeric(rules$n_findings), rules$rule_id), , drop = FALSE]
rules$.severity_order <- NULL
cat(table_html(rules))
```

## Numeric parsing quality

Parse-success percentages below are presentation-only ratios of the supplied
`n_parsed` and `n_populated` counts; underlying counts and statistics are not
recomputed.

```{r, results='asis'}
numeric <- p$numeric_summaries
numeric_display <- numeric[c("section_id", "question_id", "canonical_name", "units",
  "n_total", "n_populated", "n_parsed", "n_unparsed", "n_missing")]
numeric_display$parse_success_percent <- ifelse(
  numeric_display$n_populated > 0,
  100 * numeric_display$n_parsed / numeric_display$n_populated, NA_real_)
cat(table_html(numeric_display))
```

## Date parsing quality

```{r, results='asis'}
dates <- p$date_summaries[c("section_id", "question_id", "canonical_name",
  "n_total", "n_populated", "n_parsed", "n_unparsed", "n_missing")]
dates$parse_success_percent <- ifelse(
  dates$n_populated > 0, 100 * dates$n_parsed / dates$n_populated, NA_real_)
cat(table_html(dates))
```

## Missingness overview

Missingness percentages below are presentation-only ratios of the supplied
missing and total counts. They are descriptive displays, not quality thresholds.

```{r, results='asis'}
scalar <- p$scalar_frequencies
scalar_missing <- scalar[!duplicated(scalar$canonical_name),
  c("section_id", "question_id", "canonical_name", "n_cases_total", "n_answered", "n_missing")]
names(scalar_missing)[names(scalar_missing) == "n_cases_total"] <- "n_total"
numeric_missing <- numeric[c("section_id", "question_id", "canonical_name", "n_total", "n_missing")]
date_missing <- dates[c("section_id", "question_id", "canonical_name", "n_total", "n_missing")]
missingness <- dplyr::bind_rows(scalar_missing, numeric_missing, date_missing)
missingness$missing_percent <- ifelse(missingness$n_total > 0,
  100 * missingness$n_missing / missingness$n_total, NA_real_)
missingness <- missingness[order(-missingness$missing_percent, missingness$canonical_name), , drop = FALSE]
cat(table_html(missingness))
```

## Privacy review summary

Only aggregate privacy metadata is included below. Review rows and tested
strings are never staged for the report.

```{r, results='asis'}
if (nrow(privacy$manifest) > 0L) cat(table_html(privacy$manifest), "\n\n", sep = "") else
  cat("No de-identification manifest was available.\n\n")
cat(table_html(privacy$review), "\n\n", sep = "")
cat(table_html(privacy$leak_audit))
```

## Repeated-table completeness

```{r, results='asis'}
tables <- p$repeated_table_summaries[c("table_name", "n_cases_total",
  "n_cases_with_records", "n_records_total", "median_records_per_case",
  "max_records_per_case")]
cat(table_html(tables))
```

## Provenance

```{r, results='asis'}
provenance <- data.frame(
  Measure = c("Form version", "Privacy profile", "Summary schema version",
    "Summary status", "Cases", "bcapture package version", "Report generated"),
  Value = c(as.character(manifest$form_version[[1L]]), as.character(manifest$profile[[1L]]),
    as.character(manifest$summary_schema_version[[1L]]), as.character(status),
    as.character(manifest$n_cases[[1L]]), as.character(manifest$package_version[[1L]]),
    as.character(params$generated_at))
)
cat(table_html(provenance))
cat("\n\n<p class='epi-caption'>Generated by bcapture from de-identified Initial Epi summary products.</p>\n")
```
