---
title: "Initial Epi Descriptive Summary"
author: "bcapture"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    self-contained: true
    code-fold: false
css: assets/epi-report.css
execute:
  echo: false
  warning: false
  message: false
params:
  data_file: null
  plot_helpers: null
  report_title: "Initial Epi Descriptive Summary"
  generated_at: null
---

```{r}
data <- readRDS(params$data_file)
p <- data
if (!requireNamespace("bcapture", quietly = TRUE)) {
  if (is.null(params$plot_helpers) || !file.exists(params$plot_helpers))
    stop("bcapture package is required to render this report.")
  source(params$plot_helpers, local = .GlobalEnv)
}
plot_epi_validation_report <- if (requireNamespace("bcapture", quietly = TRUE))
  bcapture::plot_epi_validation else plot_epi_validation
plot_epi_categorical_report <- if (requireNamespace("bcapture", quietly = TRUE))
  bcapture::plot_epi_categorical else plot_epi_categorical
plot_epi_multiselect_report <- if (requireNamespace("bcapture", quietly = TRUE))
  bcapture::plot_epi_multiselect else plot_epi_multiselect
plot_epi_numeric_report <- if (requireNamespace("bcapture", quietly = TRUE))
  bcapture::plot_epi_numeric else plot_epi_numeric
overview <- p$dataset_overview
manifest <- p$summary_manifest
status <- as.character(manifest$status[[1L]])

escape_html <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  gsub('"', "&quot;", x, fixed = TRUE)
}
fmt <- function(x) ifelse(is.na(x) | !nzchar(as.character(x)), "-", as.character(x))
fmt_integer <- function(x) {
  ifelse(is.na(x), "-", format(as.integer(round(as.numeric(x))), big.mark = ",", trim = TRUE))
}
fmt_decimal <- function(x, digits = 1L) {
  ifelse(is.na(x), "-", format(round(as.numeric(x), digits), nsmall = 0L,
    trim = TRUE, scientific = FALSE))
}
fmt_percent <- function(x) fmt_decimal(x, 1L)
humanize_name <- function(x) {
  x <- gsub("_+", " ", as.character(x))
  tools::toTitleCase(x)
}
question_label <- function(x) {
  text <- if ("question_text" %in% names(x)) trimws(as.character(x$question_text)) else rep("", nrow(x))
  fallback <- if ("canonical_name" %in% names(x)) humanize_name(x$canonical_name) else rep("", nrow(x))
  use_text <- !is.na(text) & nzchar(text) & nchar(text) <= 80L
  ifelse(use_text, text, fallback)
}
section_label <- function(x) {
  text <- if ("section_name" %in% names(x)) trimws(as.character(x$section_name)) else rep("", nrow(x))
  fallback <- if ("section_id" %in% names(x)) as.character(x$section_id) else rep("", nrow(x))
  ifelse(!is.na(text) & nzchar(text), text, fallback)
}
table_html <- function(x, class = "epi-table") {
  if (nrow(x) == 0L) return("No rows to display.")
  classes <- unique(c("epi-table", strsplit(class, " +")[[1L]]))
  html <- knitr::kable(x, format = "html", escape = TRUE,
    table.attr = paste0('class="', paste(classes, collapse = " "), '"'))
  paste0('<div class="epi-table-scroll">', html, "</div>")
}
section_rows <- function(...) {
  rows <- list(...)
  ids <- unique(unlist(lapply(rows, function(x) if (nrow(x)) as.character(x$section_id) else character())))
  ids[!is.na(ids) & nzchar(ids)]
}
wrap_label <- function(x, width = 36L) {
  vapply(x, function(value) paste(strwrap(as.character(value), width = width), collapse = "\n"), character(1))
}
```

::: {.callout-note}
**`r params$report_title`**

This report describes de-identified Initial Epidemiological Interview data for
controlled analytical use. It does not contain the private re-identification
crosswalk and is not automatically suitable for unrestricted public release.
:::

```{r, results='asis'}
if (identical(status, "review")) {
  cat("::: {.callout-warning}\n**Analytical status: REVIEW**\n\nReview findings remain in the summary products; interpret this report accordingly.\n:::\n")
} else {
  cat("::: {.callout-note}\n**Analytical status: PASSED**\n\nThe summary products passed the configured analytical gates.\n:::\n")
}
```

## Dataset overview

The values below are supplied by `summarize_epi()` in `dataset_overview.csv`.

```{r, results='asis'}
overview_table <- data.frame(
  Measure = c("Cases", "Counties represented", "Interview date range",
              "Validation: valid cases", "Validation: review cases",
              "Validation: error cases", "Repeated-table records"),
  Value = c(
    overview$n_cases[[1L]], overview$n_counties[[1L]],
    paste(fmt(overview$first_interview_date[[1L]]), "to", fmt(overview$last_interview_date[[1L]])),
    overview$n_validation_valid[[1L]], overview$n_validation_review[[1L]],
    overview$n_validation_error[[1L]], overview$n_repeated_records[[1L]]
  )
)
cat(table_html(overview_table))
```

## Validation status

```{r, results='asis'}
validation <- p$validation_status_summary
print(plot_epi_validation_report(p, type = "status"))
validation_display <- data.frame(
  Status = as.character(validation$validation_status),
  Cases = fmt_integer(validation$n_cases),
  `Percent of cases` = fmt_percent(validation$percent_of_cases),
  check.names = FALSE
)
cat("\n\n", table_html(validation_display, "epi-table-compact"), sep = "")
```

## Questionnaire summaries by section

Categorical percentages and counts in this section are displayed directly from
the scalar and multiselect summary products. Percentages use the denominator
specified in each product.

```{r, results='asis'}
scalar <- p$scalar_frequencies
multi <- p$multiselect_frequencies
repeated_categorical <- p$repeated_categorical_frequencies
sections <- section_rows(scalar, multi, repeated_categorical)
for (section_id in sections) {
  section_name <- c(
    as.character(scalar$section_name[scalar$section_id == section_id]),
    as.character(multi$section_name[multi$section_id == section_id]),
    as.character(repeated_categorical$section_name[repeated_categorical$section_id == section_id])
  )
  section_name <- section_name[!is.na(section_name) & nzchar(section_name)][1L]
  cat("### Section ", escape_html(section_id),
      if (!is.na(section_name)) paste0(": ", escape_html(section_name)) else "", "\n\n", sep = "")
  section_scalar <- scalar[scalar$section_id == section_id, , drop = FALSE]
  for (question_id in unique(as.character(section_scalar$question_id))) {
    current <- section_scalar[section_scalar$question_id == question_id, , drop = FALSE]
    question_text <- question_label(current)[[1L]]
    cat("#### ", escape_html(question_id), ": ", escape_html(question_text), "\n\n", sep = "")
    current$response_label <- fmt(current$response_label)
    if (all(as.numeric(current$n_answered) == 0)) {
      cat("No responses recorded. Missing: ", current$n_missing[[1L]],
          " of ", current$n_cases_total[[1L]], " cases.\n\n", sep = "")
      next
    }
    if (nrow(current) <= 8L) {
      print(plot_epi_categorical_report(p, variable = unique(as.character(current$canonical_name))[[1L]],
        denominator = "answered", show_counts = FALSE, show_missing = FALSE))
    } else {
      display <- current[c("response_label", "n_response", "percent_of_answered", "n_answered", "n_missing", "n_cases_total")]
      names(display) <- c("Response", "Responses", "Percent of answered", "Answered", "Missing", "Cases")
      cat(table_html(display), "\n\n", sep = "")
    }
    cat("<p class='epi-caption'>Response categories: ",
        paste(vapply(as.character(current$response_label), escape_html, character(1)),
          collapse = ", "), ".</p>\n", sep = "")
    cat("<p class='epi-caption'>Answered: ", current$n_answered[[1L]], " of ",
        current$n_cases_total[[1L]], " cases; missing: ", current$n_missing[[1L]],
        ". Zero-count codebook categories are retained.</p>\n", sep = "")
  }
  section_multi <- multi[multi$section_id == section_id, , drop = FALSE]
  for (question_id in unique(as.character(section_multi$question_id))) {
    current <- section_multi[section_multi$question_id == question_id, , drop = FALSE]
    cat("#### ", escape_html(question_id), ": ", escape_html(question_label(current)[[1L]]), "\n\n", sep = "")
    print(plot_epi_multiselect_report(p, variable = question_id,
      denominator = "all_cases"))
    cat("<p class='epi-caption'>Cases with any selection: ",
        current$n_cases_with_any_selection[[1L]],
        ". Percentages within a multiselect question need not sum to 100%.</p>\n", sep = "")
  }
}
```

## Repeated-table overview

The repeated-table products retain all registered tables, including tables with
zero records. Categorical percentages below use records with a nonmissing
response, not cases.

```{r, results='asis'}
tables <- p$repeated_table_summaries
display <- tables[c("table_name", "n_cases_with_records", "n_records_total",
                    "median_records_per_case", "max_records_per_case")]
names(display) <- c("Table", "Cases", "Records", "Median / case", "Max / case")
display$Cases <- fmt_integer(display$Cases)
display$Records <- fmt_integer(display$Records)
display$`Median / case` <- fmt_decimal(display$`Median / case`)
display$`Max / case` <- fmt_integer(display$`Max / case`)
cat(table_html(display, "epi-table-compact"), "\n\n", sep = "")
cat("<p class='epi-caption'>Cases and records refer to repeated-table records; median and maximum are records per case.</p>\n\n")
if (nrow(repeated_categorical) > 0L) {
  repeated_display <- repeated_categorical[c(
    "table_name", "column_name", "response_label", "n_value", "percent_of_nonmissing")]
  repeated_display$column_name <- humanize_name(repeated_display$column_name)
  repeated_display$n_value <- fmt_integer(repeated_display$n_value)
  repeated_display$percent_of_nonmissing <- fmt_percent(repeated_display$percent_of_nonmissing)
  names(repeated_display) <- c("Table", "Variable", "Response", "N", "Percent (%)")
  cat(table_html(repeated_display, "epi-table-wide"))
}
```

## Numeric summaries

These are descriptive summaries supplied by `numeric_summaries.csv`. The
report does not reconstruct distributions from aggregate statistics.

```{r, results='asis'}
numeric <- p$numeric_summaries
numeric_display <- data.frame(
  Section = section_label(numeric),
  `Question / Variable` = question_label(numeric),
  Units = fmt(numeric$units),
  Parsed = fmt_integer(numeric$n_parsed),
  Unparsed = fmt_integer(numeric$n_unparsed),
  Missing = fmt_integer(numeric$n_missing),
  Median = fmt_decimal(numeric$median),
  Q1 = fmt_decimal(numeric$q25),
  Q3 = fmt_decimal(numeric$q75),
  Min = fmt_decimal(numeric$min),
  Max = fmt_decimal(numeric$max),
  check.names = FALSE
)
cat(table_html(numeric_display, "epi-table-wide"), "\n\n", sep = "")
plot_data <- numeric[!is.na(numeric$median), , drop = FALSE]
if (nrow(plot_data) > 0L) for (variable in unique(as.character(plot_data$canonical_name)))
  print(plot_epi_numeric_report(p, variable = variable))
```

## Date summaries

Date ranges and parsing counts are displayed directly from `date_summaries.csv`.

```{r, results='asis'}
dates <- p$date_summaries
date_display <- data.frame(
  Section = section_label(dates),
  `Question / Variable` = question_label(dates),
  Parsed = fmt_integer(dates$n_parsed),
  Unparsed = fmt_integer(dates$n_unparsed),
  Missing = fmt_integer(dates$n_missing),
  Earliest = fmt(dates$min_date),
  Median = fmt(dates$median_date),
  Latest = fmt(dates$max_date),
  check.names = FALSE
)
cat(table_html(date_display, "epi-table-wide"))
```

## Notes on denominators and missingness

Categorical summaries distinguish answered cases from missing cases. Explicit
answered categories such as `Not applicable` remain response categories. Scalar
categorical percentages use `percent_of_answered`; multiselect percentages show
their denominator in the axis and caption; repeated categorical percentages use
records with a nonmissing response.

## Provenance

```{r, results='asis'}
provenance <- data.frame(
  Measure = c("Form version", "Privacy profile", "Summary schema version",
              "Summary status", "Cases", "bcapture package version", "Report generated"),
  Value = c(as.character(manifest$form_version[[1L]]), as.character(manifest$profile[[1L]]),
    as.character(manifest$summary_schema_version[[1L]]), as.character(status),
    as.character(manifest$n_cases[[1L]]), as.character(manifest$package_version[[1L]]),
    as.character(params$generated_at))
)
cat(table_html(provenance))
cat("\n\n<p class='epi-caption'>Generated by bcapture from de-identified Initial Epi summary products.</p>\n")
```
