---
title: "Initial Epi Analytical Feature Review"
author: "bcapture"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    self-contained: true
    code-fold: false
css: assets/epi-report.css
execute:
  echo: false
  warning: false
  message: false
params:
  data_file: null
  plot_helpers: null
  report_title: "Initial Epi Analytical Feature Review"
  generated_at: null
---

```{r}
p <- readRDS(params$data_file)
manifest <- p$feature_summary_manifest
overview <- p$overview
status <- as.character(manifest$status[[1L]])

fmt_integer <- function(x) {
  ifelse(is.na(x), "-", format(as.integer(round(as.numeric(x))), big.mark = ",", trim = TRUE))
}
fmt_decimal <- function(x, digits = 1L) {
  ifelse(is.na(x), "-", format(round(as.numeric(x), digits), nsmall = 0L,
    trim = TRUE, scientific = FALSE))
}
fmt_percent <- function(x) paste0(fmt_decimal(x, 1L), "%")
fmt <- function(x) ifelse(is.na(x) | !nzchar(as.character(x)), "-", as.character(x))
humanize <- function(x) tools::toTitleCase(gsub("_+", " ", as.character(x)))
wrap_label <- function(x, width = 42L) vapply(as.character(x), function(value) {
  paste(strwrap(value, width = width), collapse = "\n")
}, character(1))
table_html <- function(x, class = "epi-table") {
  if (nrow(x) == 0L) return("No rows to display.")
  classes <- unique(c("epi-table", strsplit(class, " +")[[1L]]))
  html <- knitr::kable(x, format = "html", escape = TRUE,
    table.attr = paste0('class="', paste(classes, collapse = " "), '"'))
  paste0('<div class="epi-table-scroll">', html, "</div>")
}
theme_feature <- function() {
  ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major.y = ggplot2::element_blank(),
      plot.title = ggplot2::element_text(face = "bold"),
      plot.caption = ggplot2::element_text(hjust = 0, colour = "#52616b"),
      axis.text.y = ggplot2::element_text(lineheight = 0.9)
    )
}
```

::: {.callout-note}
**Controlled-use notice**

This report summarizes de-identified case-level analytical features derived
from structured Initial Epi questionnaire data. These features are descriptive
representations of questionnaire observations. They are not risk scores and do
not imply causation.

This report does not contain a private re-identification crosswalk and is not
automatically suitable for unrestricted public release.
:::

```{r, results='asis'}
if (identical(status, "review")) {
  cat("::: {.callout-warning}\n**Analytical status: REVIEW**\n\nFeature or consistency findings remain for review. No inferential conclusion is generated.\n:::\n")
} else {
  cat("::: {.callout-note}\n**Analytical status: PASSED**\n\nThe descriptive feature summaries passed their configured analytical gates.\n:::\n")
}
```

## Feature overview

All aggregate statistics in this report are read from `features/summary/*.csv`.
The report does not recompute feature derivations.

```{r, results='asis'}
overview_table <- data.frame(
  Measure = c("Cases", "Features", "Domains", "Binary features", "Categorical features",
    "Numeric features", "Count features", "Consistency findings",
    "Consistency warnings", "Consistency errors", "Source feature status"),
  Value = c(overview$n_cases, overview$n_features, overview$n_domains,
    overview$n_binary, overview$n_categorical, overview$n_numeric, overview$n_count,
    overview$n_consistency_findings, overview$n_consistency_warnings,
    overview$n_consistency_errors, overview$source_feature_status)
)
cat(table_html(overview_table, "epi-table-compact"))
```

## Domain coverage

Coverage describes the completeness of recorded feature information. It is not
a measure of domain importance, quality, or risk.

```{r, fig.height=5.8, fig.width=9}
coverage <- p$domain_coverage
coverage <- coverage[order(coverage$percent_known_cells), , drop = FALSE]
coverage$display_domain <- factor(coverage$domain_label, levels = coverage$domain_label)
ggplot2::ggplot(coverage, ggplot2::aes(x = display_domain, y = percent_known_cells)) +
  ggplot2::geom_col(fill = "#6b8e9e") + ggplot2::coord_flip() +
  ggplot2::scale_x_discrete(labels = function(x) wrap_label(x, 42L)) +
  ggplot2::labs(x = NULL, y = "Percent of case-feature cells known",
    caption = "Information coverage only; a TRUE state is not treated as better or worse.") +
  theme_feature()
```

```{r, results='asis'}
coverage_table <- data.frame(
  Domain = coverage$domain_label,
  Features = fmt_integer(coverage$n_features),
  Cases = fmt_integer(coverage$n_cases),
  `Known cells` = fmt_integer(coverage$n_known_cells),
  `Missing cells` = fmt_integer(coverage$n_missing_cells),
  `Known cells (%)` = fmt_percent(coverage$percent_known_cells),
  `Cases with any known (%)` = fmt_percent(coverage$percent_cases_with_any_known_feature),
  `Cases with all known (%)` = fmt_percent(coverage$percent_cases_with_all_features_known),
  check.names = FALSE
)
cat(table_html(coverage_table))
```

## Binary feature prevalence

Percentages use cases with an explicit TRUE or FALSE derived value as the
denominator. Missing feature values are not treated as FALSE.

```{r, fig.height=8.5, fig.width=10}
binary <- p$binary[order(p$binary$percent_true_of_known), , drop = FALSE]
binary$display_feature <- factor(binary$feature_label, levels = binary$feature_label)
ggplot2::ggplot(binary, ggplot2::aes(x = display_feature, y = percent_true_of_known)) +
  ggplot2::geom_col(fill = "#2c6e8f") + ggplot2::coord_flip() +
  ggplot2::scale_x_discrete(labels = function(x) wrap_label(x, 48L)) +
  ggplot2::labs(x = NULL, y = "Percent TRUE among known responses",
    caption = "Descriptive prevalence of recorded derived states; no risk direction is assigned.") +
  theme_feature()
```

```{r, results='asis'}
binary_table <- data.frame(
  Feature = binary$feature_label, Domain = binary$domain_label,
  Known = fmt_integer(binary$n_known), Missing = fmt_integer(binary$n_missing),
  `TRUE among known (%)` = fmt_percent(binary$percent_true_of_known),
  `TRUE among all cases (%)` = fmt_percent(binary$percent_true_of_all_cases),
  check.names = FALSE
)
cat(table_html(binary_table))
```

## Categorical feature distributions

`Don't know` and `Not applicable` remain substantive questionnaire states and
are included in the known-state denominator. Registry-defined categories with
zero observations remain visible.

```{r, results='asis'}
categorical <- p$categorical
for (feature_name in unique(categorical$feature_name)) {
  current <- categorical[categorical$feature_name == feature_name, , drop = FALSE]
  cat("### ", current$feature_label[[1L]], "\n\n", sep = "")
  display <- data.frame(
    Category = current$category, Recorded = fmt_integer(current$n),
    `Percent of known states` = fmt_percent(current$percent_of_known),
    `Percent of all cases` = fmt_percent(current$percent_of_all_cases),
    Missing = fmt_integer(current$n_missing), check.names = FALSE
  )
  cat(table_html(display, "epi-table-compact"), "\n\n")
}
```

## Count and numeric features

Aggregate descriptive statistics below come from the feature summary products.
For the figures only, the renderer stages the minimum necessary de-identified
numeric/count values, without case identifiers. This is the sole case-level
data exception in this aggregate report; no fitted curves or inferential
statistics are added.

For count features, zero means zero qualifying retained-content records. It
does not confirm absence of the underlying event.

```{r, results='asis'}
numeric <- p$numeric
numeric_table <- data.frame(
  Feature = numeric$feature_label, Domain = numeric$domain_label, Unit = fmt(numeric$unit),
  Recorded = fmt_integer(numeric$n_nonmissing), Missing = fmt_integer(numeric$n_missing),
  Mean = fmt_decimal(numeric$mean, 2L), SD = fmt_decimal(numeric$sd, 2L),
  Median = fmt_decimal(numeric$median, 2L), Q25 = fmt_decimal(numeric$q25, 2L),
  Q75 = fmt_decimal(numeric$q75, 2L), Minimum = fmt_decimal(numeric$min, 2L),
  Maximum = fmt_decimal(numeric$max, 2L), check.names = FALSE
)
cat("### Numeric summaries\n\n", table_html(numeric_table), "\n\n", sep = "")

counts <- p$count
count_table <- data.frame(
  Feature = counts$feature_label, Domain = counts$domain_label,
  Recorded = fmt_integer(counts$n_nonmissing), Zero = fmt_integer(counts$n_zero),
  Positive = fmt_integer(counts$n_positive), Mean = fmt_decimal(counts$mean, 2L),
  SD = fmt_decimal(counts$sd, 2L), Median = fmt_decimal(counts$median, 2L),
  Q25 = fmt_decimal(counts$q25, 2L), Q75 = fmt_decimal(counts$q75, 2L),
  Minimum = fmt_decimal(counts$min, 2L), Maximum = fmt_decimal(counts$max, 2L),
  check.names = FALSE
)
cat("### Count summaries\n\n", table_html(count_table), "\n\n", sep = "")
```

```{r, fig.height=3.1, fig.width=9, results='asis'}
distribution <- p$distribution_data
for (feature_name in unique(distribution$feature_name)) {
  current <- distribution[distribution$feature_name == feature_name, , drop = FALSE]
  current$display_feature <- factor(wrap_label(current$feature_label, 48L))
  plot <- ggplot2::ggplot(current, ggplot2::aes(x = display_feature, y = numeric_value)) +
    ggplot2::geom_boxplot(width = 0.28, outlier.shape = NA, fill = "#d8e4e9",
      colour = "#2c6e8f") +
    ggplot2::geom_point(position = ggplot2::position_jitter(
      width = 0.09, height = 0, seed = 1729L), alpha = 0.55, size = 1.8,
      colour = "#24313a") + ggplot2::coord_flip() +
    ggplot2::labs(title = current$feature_label[[1L]], x = NULL,
      y = if (current$feature_type[[1L]] == "count") "Recorded qualifying rows" else "Recorded value") +
    theme_feature()
  print(plot)
}
```

## Feature missingness

```{r, fig.height=10.5, fig.width=10}
missingness <- p$missingness[order(p$missingness$percent_missing, decreasing = TRUE), , drop = FALSE]
missingness$display_feature <- factor(missingness$feature_label,
  levels = rev(missingness$feature_label))
ggplot2::ggplot(missingness, ggplot2::aes(x = display_feature, y = percent_missing)) +
  ggplot2::geom_col(fill = "#7b8790") + ggplot2::coord_flip() +
  ggplot2::scale_x_discrete(labels = function(x) wrap_label(x, 48L)) +
  ggplot2::labs(x = NULL, y = "Percent missing",
    caption = "Don't know and Not applicable are recorded categorical states, not missing values.") +
  theme_feature()
```

## Consistency findings

Consistency findings preserve questionnaire and child-record discrepancies for
review. Counts do not expose case identifiers or source values.

```{r, results='asis'}
severity <- p$consistency[p$consistency$summary_level == "severity_total", , drop = FALSE]
severity_display <- data.frame(
  Severity = severity$severity, Findings = fmt_integer(severity$n_findings),
  `Cases affected` = fmt_integer(severity$n_cases), check.names = FALSE
)
cat(table_html(severity_display, "epi-table-compact"))
```

```{r, fig.height=5.5, fig.width=10}
findings <- p$consistency[p$consistency$summary_level == "finding", , drop = FALSE]
if (nrow(findings) == 0L) {
  plot.new(); text(0.5, 0.5, "No consistency findings recorded")
} else {
  key <- paste(findings$domain_label, findings$finding_type, findings$severity, sep = "\r")
  sums <- rowsum(as.numeric(findings$n_findings), key, reorder = FALSE)
  parts <- strsplit(rownames(sums), "\r", fixed = TRUE)
  display <- data.frame(
    Domain = vapply(parts, `[[`, character(1), 1L),
    Finding = vapply(parts, `[[`, character(1), 2L),
    Severity = vapply(parts, `[[`, character(1), 3L),
    Count = as.numeric(sums[, 1L]), stringsAsFactors = FALSE
  )
  display$Label <- wrap_label(paste0(display$Domain, ": ", humanize(display$Finding)), 52L)
  display$Severity <- factor(display$Severity, levels = c("ERROR", "WARNING", "INFO"))
  print(ggplot2::ggplot(display, ggplot2::aes(x = Label, y = Count, fill = Severity)) +
    ggplot2::geom_col() + ggplot2::coord_flip() +
    ggplot2::scale_fill_manual(values = c(
      ERROR = "#a94a4a", WARNING = "#9a6b20", INFO = "#557d8b"), drop = FALSE) +
    ggplot2::labs(x = "Domain and finding type", y = "Findings") + theme_feature())
}
```

## Feature dictionary and definitions

Features are grouped using the domain labels supplied by the versioned feature
dictionary. Derivation code and case-level values are not displayed.

```{r, results='asis'}
dictionary <- p$feature_dictionary
for (domain_id in unique(dictionary$domain_id)) {
  current <- dictionary[dictionary$domain_id == domain_id, , drop = FALSE]
  cat("### ", current$domain_label[[1L]], "\n\n", sep = "")
  source <- ifelse(!is.na(current$source_question_id) & nzchar(current$source_question_id),
    paste0("Question ", current$source_question_id), humanize(current$source_table))
  display <- data.frame(
    Feature = humanize(current$feature_name), Label = current$feature_label,
    Type = humanize(current$feature_type), Source = source,
    Description = current$description, check.names = FALSE
  )
  cat(table_html(display), "\n\n")
}
```

## Provenance

```{r, results='asis'}
provenance <- data.frame(
  Field = c("Feature summary schema", "Feature version", "Form version",
    "Feature registry hash", "Source feature manifest status", "Summary status",
    "bcapture version", "Generated"),
  Value = c(manifest$feature_summary_schema_version, manifest$feature_version,
    manifest$form_version, manifest$feature_registry_hash,
    manifest$source_feature_manifest_status, manifest$status,
    manifest$package_version, params$generated_at)
)
cat(table_html(provenance, "epi-table-compact"))
```

## Scope boundary and next phase

This review is descriptive only. It contains no regression, odds ratio,
relative risk, p-value, confidence interval, feature importance, machine
learning, composite score, risk weight, or causal interpretation.

Before any future modeling, the next phase must begin with an explicit outcome
definition and contracts for the unit of analysis, sampling frame, candidate
predictors, potentially post-outcome or endogenous features, and temporal
ordering.
