---
title: "BCAP workflow: extraction through analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BCAP workflow: extraction through analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(bcapture)

example_pdf <- system.file(
  "extdata", "examples", "synthetic_bcap_audit_print.pdf",
  package = "bcapture"
)
stopifnot(nzchar(example_pdf), file.exists(example_pdf))
```

This tutorial uses a fictional BCAP audit PDF included with the package. The
values are synthetic and must not be replaced with completed production forms
in a package installation or source repository.

## Keep confidential, safe, crosswalk, and analysis data separate

The BCAP workflow has four operational layers:

1. **Confidential extraction output** retains source-form values and source
   provenance for controlled processing.
2. **Safe/de-identified output** contains only fields permitted by the
   versioned BCAP privacy policy.
3. **Private crosswalk** stores mappings needed for authorized
   re-identification and must remain outside the safe output.
4. **Analysis-ready output** contains the safe BCAP records, repeated-item
   product, provenance, metadata, QA, and manifests.

The safe output and the private crosswalk are separate by construction. Do not
commit either production directory to a repository.

## Extract and validate

This opt-in example extracts the synthetic PDF into a temporary directory and
then runs strict BCAP validation. PDF extraction requires the optional Python
PDF dependencies, so the chunk is not evaluated during vignette builds.

```{r bcap-extract-validate, eval=FALSE}
input_dir <- tempfile("bcapture-bcap-input-")
dir.create(input_dir, recursive = TRUE)
file.copy(example_pdf, file.path(input_dir, "synthetic_bcap_audit.pdf"))

extraction_dir <- tempfile("bcapture-bcap-extraction-")
extraction_manifest <- extract_hpai(
  in_dir = input_dir,
  out_dir = extraction_dir,
  diagnostics = TRUE,
  quiet = TRUE
)

validation <- validate_hpai(
  out_dir = extraction_dir,
  write = TRUE,
  strict = TRUE,
  quiet = TRUE
)
validation$status
```

`validate_hpai()` is a semantic gate. It does not repair extracted values.
Unsupported fields, invalid controls, and unparseable dates remain visible as
findings and prevent strict finalization.

## De-identify, privacy-audit, and collate

`deidentify_hpai()` requires strict validation and writes the safe output and
private crosswalk to distinct destinations. A second privacy audit can be run
against the safe output before Phase 3 collation.

```{r bcap-deidentify-collate, eval=FALSE}
safe_dir <- tempfile("bcapture-bcap-safe-")
private_crosswalk_dir <- tempfile("bcapture-bcap-crosswalk-")

safe <- deidentify_hpai(
  out_dir = extraction_dir,
  deidentified_dir = safe_dir,
  crosswalk_dir = private_crosswalk_dir,
  strict = TRUE,
  quiet = TRUE
)

privacy <- validate_hpai_privacy(
  deidentified_dir = safe_dir,
  crosswalk_dir = private_crosswalk_dir,
  strict = TRUE,
  write = TRUE,
  quiet = TRUE
)

analysis_dir <- tempfile("bcapture-bcap-analysis-")
analysis <- collate_bcap(
  bcap_dir = safe_dir,
  output_dir = analysis_dir,
  bcap_crosswalk_dir = private_crosswalk_dir,
  run_id = "tutorial-bcap",
  quiet = TRUE
)

analysis$status
list.files(analysis_dir)
```

The BCAP analysis family is independent of Initial Epi. Its core products are
`records`, `repeated_items`, `provenance`, and `field_metadata`. The current
BCAP template has no relational repeated structure, so `repeated_items` is an
explicit empty typed table rather than an omitted product.

## Inspect field metadata

Use the versioned metadata dictionary to understand how source-form fields map
to safe analysis output.

```{r bcap-metadata}
bcap_metadata <- field_metadata("bcap")
bcap_metadata[
  bcap_metadata$field_name %in% c("premises_address", "npin", "Q1_1_1"),
  c("field_name", "canonical_name", "question_text", "section",
    "question_number", "source_page", "response_type", "privacy_class",
    "included_in_safe_output")
]
```

Withheld fields remain in metadata even though they are absent from
analysis-ready `records`. This preserves source-question traceability without
exposing withheld values.

## Fail closed and keep cross-reference optional

Do not continue after a failed validation or privacy audit:

```{r bcap-fail-closed, eval=FALSE}
if (!identical(validation$status, "passed")) {
  stop("BCAP validation failed; no safe output should be finalized.")
}
if (!identical(privacy$status, "passed")) {
  stop("BCAP privacy audit failed; review findings before collation.")
}
```

BCAP can be analyzed without Initial Epi. If both safe families and both
private crosswalks are available, `cross_reference_hpai()` may be called as a
separate optional step. It uses only exact authoritative premises keys; it
does not infer links from dates, names, geography, or response similarity.
Zero links is a valid outcome, and the Epi and BCAP output families remain
separate.
