---
title: "Initial Epi workflow: extraction through analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Initial Epi workflow: extraction through analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(bcapture)

example_pdf <- system.file(
  "extdata", "examples", "synthetic_initial_epi_print.pdf",
  package = "bcapture"
)
stopifnot(nzchar(example_pdf), file.exists(example_pdf))
```

This tutorial uses the fictional PDF shipped with the package. It contains no
production records or real identifiers. Replace the temporary input and output
directories below with controlled locations when adapting the workflow to an
approved dataset.

## Keep the four data layers separate

The supported workflow has four distinct layers:

1. **Confidential extraction output** contains source-form values and source
   provenance. Treat it as restricted working data.
2. **Safe/de-identified output** contains pseudonyms or normalized values after
   semantic validation and privacy checks pass.
3. **Private crosswalk** contains re-identification and source-linkage mappings.
   It is created outside the safe output and must never be distributed with it.
4. **Analysis-ready output** contains the separate Epi family products used by
   downstream analysis. It does not contain the private crosswalk.

The analysis-ready Phase 3 API is intentionally separate from the existing
`collate_epi()` API. `collate_epi()` converts raw extraction products into
semantic relational products; `collate_epi_analysis()` consumes the validated
safe output and writes the final Epi analysis family.

## Extract, collate, and validate

The following opt-in example copies the synthetic PDF into a temporary input
directory, extracts it, performs the established semantic collation, and runs
validation. Extraction of PDFs requires the optional Python PDF dependencies;
the code is therefore shown but not evaluated while the vignette is built.

```{r epi-extract-validate, eval=FALSE}
input_dir <- tempfile("bcapture-epi-input-")
dir.create(input_dir, recursive = TRUE)
file.copy(example_pdf, file.path(input_dir, "synthetic_initial_epi.pdf"))

extraction_dir <- tempfile("bcapture-epi-extraction-")
extraction_manifest <- extract_epi(
  in_dir = input_dir,
  out_dir = extraction_dir,
  diagnostics = TRUE,
  quiet = TRUE
)

# Convert extraction rows into the semantic Epi products required by validation.
collated <- collate_epi(
  out_dir = extraction_dir,
  quiet = TRUE
)

validation <- validate_epi(
  out_dir = extraction_dir,
  write = TRUE,
  strict = FALSE,
  quiet = TRUE
)
validation
```

Validation reports findings; it does not repair or reinterpret source values.
For a production run, review the findings and continue only when the intended
privacy and semantic gates are satisfied.

## De-identify and finalize the Epi analysis family

`deidentify_epi()` writes a safe output tree and a physically separate private
crosswalk. The Phase 3 finalizer then consumes only the safe tree.

```{r epi-deidentify-collate, eval=FALSE}
safe_dir <- tempfile("bcapture-epi-safe-")
private_crosswalk_dir <- tempfile("bcapture-epi-crosswalk-")

safe <- deidentify_epi(
  out_dir = extraction_dir,
  deidentified_dir = safe_dir,
  crosswalk_dir = private_crosswalk_dir,
  strict = TRUE,
  quiet = TRUE
)

analysis_dir <- tempfile("bcapture-epi-analysis-")
analysis <- collate_epi_analysis(
  epi_dir = safe_dir,
  output_dir = analysis_dir,
  epi_crosswalk_dir = private_crosswalk_dir,
  run_id = "tutorial-epi",
  quiet = TRUE
)

analysis$status
list.files(analysis_dir)
```

The resulting Epi family contains `records`, `repeated_items`, `provenance`,
and `field_metadata`, together with QA and manifest products. Repeated Epi
tables remain long rather than being flattened into the parent record table.

## Inspect field metadata

Field metadata is generated from the version-controlled dictionary and privacy
policy, not from production values.

```{r epi-metadata}
epi_metadata <- field_metadata("epi")
epi_metadata[
  epi_metadata$field_name %in% c("premises_address", "premises_id"),
  c("field_name", "canonical_name", "question_text", "section",
    "question_number", "source_page", "response_type", "privacy_class",
    "included_in_safe_output")
]
```

Metadata includes fields that are intentionally withheld from safe output.
This lets analysts understand the source-form question and its privacy
disposition without receiving the withheld value.

## Fail-closed behavior and optional linkage

If validation or privacy checks fail, `deidentify_epi()` stops before a safe
output is finalized. For example, a production caller should treat the
following as a blocking condition rather than bypassing it:

```{r epi-fail-closed, eval=FALSE}
if (!identical(validation$status, "passed")) {
  stop("Review validation findings before de-identification.")
}
```

Cross-form linkage is not required for Epi analysis. When both safe families
are available, `cross_reference_hpai()` is an optional, fail-closed utility
that uses only exact authoritative premises keys from both private crosswalks.
It never infers a relationship from dates, names, geography, or similarity;
zero authoritative links is valid.
